Data Management for LSST Special Programs

### 2023-11-09 sec_proc.tex
fix program/WFD terms

Recall that the term ``regular processing" is used only within this document,
and refers to the image processing pipelines that are designed for, 
and will be applied to, the WFD program's observations (Section~\ref{sec:intro}).

Decisions about when to apply regular processing to Special Programs data,
or when to include it in the data products for the WFD program 
(e.g., if it improves the all-sky coadds), 
are ultimately left to the discretion of the Rubin Observatory's 
Data Management and System Performance teams.

@@

\textbf{How WFD and Special Program data would co-exist.}
No ``unique and separate" data products for the Special Progams would be 
produced by regular Prompt processing.
Special Programs data that is processed by the Prompt pipeline would 
contribute to the Prompt data products for the WFD program as 
described in Section 3 of the DPDD \citedsp{lse-163}. 
These data products are the results of Difference Image Analysis (DIA),
such as the difference images, catalogs of sources detected in difference
images ({\tt DiaSources}) and associated static-sky {\tt DiaSources}
into {\tt DiaObjects}, and alert packets.
Including visits from Special Programs in regular Prompt processing alongside
vists from the WFD program is not, in general, anticipated to affect WFD science goals.
For example, analyses for a WFD-only subset could still be done using the program and
region labels described in Section~\ref{ssec:sci_labels}, which would be
propogated to difference images, difference-image catalogss, and alerts.

@@

\item \textbf{DIA Object histories may become too large for the sizing model.}
Alert packets contain the full records of all associated 
{\tt DiaSources} from the past 12 months, but the alert
stream bandwidth is sized for the expected histories for
WFD program fields. 
The Prompt pipeline resources are also sized for the
WFD program, and it might not be possible to load up
thousands of epochs at a time.
The Data Management team will have to test the realized
capabilities of Prompt processing and alert distribtuion,
and potentially impose a mitigation strategy such as
limiting histories to the last $N$ observations instead
of the last 12 months in heavily-observed regions.
Alerts from Special Programs data are subject to the same latency
requirements as WFD program alerts (Section~\ref{ssec:req_proc}).

@@

These two potential issues pose challenges, but are not necessarily showstoppers in processing 
Special Programs data with the Prompt pipelines. 
The overall impact on time-domain science would still be positive, even 
if mitigations are needed for these issues.
For example, brokers and users would be able to use the region and program labels
in the data as context (i.e., as flags) and avoid including 
limited-history or potentially-compromised {\tt DiaSources} in their
analyses if necessary.

@@

\textbf{Time-domain DIA data products --}
This includes the results of the annual reprocessing of WFD program data with 
Difference Image Analysis (DIA), and the production of Data Release
versions of the processed images (single-visit and difference images)
and associated catalogs ({\tt DiaSource}, {\tt DiaObject}, {\tt Source},
{\tt ForcedSource}, {\tt DiaForcedSource}, and so on).
These data products will be used primarily for time-domain science.
They should include Special Programs data for the same reasons as
provided in Section~\ref{ssec:sci_pproc}, and with the same 
considerations as discussed in Section~\ref{sssec:proc_reg_prompt}.
This processing should use the same template image for a given field.

\textbf{Static-sky data products --}
This includes the tessellation and coaddition of WFD program images
and the associated multi-band {\tt Object} catalog and survey property maps.
Whether and how to include any Special Programs data in these data products
is left entirely to the discretion of the Rubin Data Management team in Operations.
As an example, perhaps Special Programs images will only be included when they 
assist with uniformity or suppress edge effects or low-order modes in the 
all-sky photometric solutions.

@@

\textbf{Joinable tables --}
Any tables for the unique and separate data products should be joinable to the 
data products for the WFD program, when possible.
This is a requirement (Section~\ref{ssec:req_dp}).

\textbf{Survey property maps --}
As for the tesslated all-sky coadded images made from WFD program data, survey property maps
should be made individually as part of any Special Programs that generates
tesselated coadded images. This is \emph{not} a requirement (Section~\ref{ssec:req_dp}).

@@

\item \textbf{Possible but not necessary (secondary science goals):}
a time-domain mini-survey (or DDF) has a secondary science goal of detecting
precursor outbursts for transients, which requires coadding images
within windows of days, weeks, and months to reach various depths.
This set of custom coadds may be considered as ``not necessary" and requiring 
User-Generated Processing.
Note that for the WFD program data, such custom coadds are also considered 
as beyond scope and in need of User-Generated Processing.

@@

Consider a Special Program in which a special region of sky is observed with standard
visits but a a special strategy or cadence which is significantly distinct from the WFD program,
and lasts for a limited amount of time.
For example, a short-term survey of the Magellanic Clouds, with dual primary science
goals in time-domain and static-sky science.

>>>>

Recall that the term ``regular processing" is used only within this document,
and refers to the image processing pipelines that are designed for, 
and will be applied to, the Main Survey's observations (Section~\ref{sec:intro}).

Decisions about when to apply regular processing to Special Programs data,
or when to include it in the data products for the Main Survey
(e.g., if it improves the all-sky coadds), 
are ultimately left to the discretion of the Rubin Observatory's 
Data Management and System Performance teams.

@@

\textbf{How Main Survey and Special Program data would co-exist.}
No ``unique and separate" data products for the Special Progams would be 
produced by regular Prompt processing.
Special Programs data that is processed by the Prompt pipeline would 
contribute to the Prompt data products for the Main Survey as 
described in Section 3 of the DPDD \citedsp{lse-163}. 
These data products are the results of Difference Image Analysis (DIA),
such as the difference images, catalogs of sources detected in difference
images ({\tt DiaSources}) and associated static-sky {\tt DiaSources}
into {\tt DiaObjects}, and alert packets.
Including visits from Special Programs in regular Prompt processing alongside
vists from the Main Survey is not, in general, anticipated to affect Main Survey 
or WFD-specific science goals.
For example, analyses for a WFD-only subset could still be done using the observing mode
and region labels described in Section~\ref{ssec:sci_labels}, which would be
propogated to difference images, difference-image catalogss, and alerts.

@@

\item \textbf{DIA Object histories may become too large for the sizing model.}
Alert packets contain the full records of all associated 
{\tt DiaSources} from the past 12 months, but the alert
stream bandwidth is sized for the expected histories for
Main Survey fields. 
The Prompt pipeline resources are also sized for the
Main Survey, and it might not be possible to load up
thousands of epochs at a time.
The Data Management team will have to test the realized
capabilities of Prompt processing and alert distribtuion,
and potentially impose a mitigation strategy such as
limiting histories to the last $N$ observations instead
of the last 12 months in heavily-observed regions.
Alerts from Special Programs data are subject to the same latency
requirements as Main Survey alerts (Section~\ref{ssec:req_proc}).

@@

These two potential issues pose challenges, but are not necessarily showstoppers 
in processing Special Programs data with the Prompt pipelines. 
The overall impact on time-domain science would still be positive, even 
if mitigations are needed for these issues.
For example, brokers and users would be able to use the region and observing mode 
labels in the data as context (i.e., as flags) and avoid including 
limited-history or potentially-compromised {\tt DiaSources} in their
analyses if necessary.

@@

\textbf{Time-domain DIA data products --}
This includes the results of the annual reprocessing of Main Survey data with 
Difference Image Analysis (DIA), and the production of Data Release
versions of the processed images (single-visit and difference images)
and associated catalogs ({\tt DiaSource}, {\tt DiaObject}, {\tt Source},
{\tt ForcedSource}, {\tt DiaForcedSource}, and so on).
These data products will be used primarily for time-domain science.
They should include Special Programs data for the same reasons as
provided in Section~\ref{ssec:sci_pproc}, and with the same 
considerations as discussed in Section~\ref{sssec:proc_reg_prompt}.
This processing should use the same template image for a given field.

\textbf{Static-sky data products --}
This includes the tessellation and coaddition of Main Survey images
and the associated multi-band {\tt Object} catalog and survey property maps.
Whether and how to include any Special Programs data in these data products
is left entirely to the discretion of the Rubin Data Management team in Operations.
As an example, perhaps Special Programs images will only be included when they 
assist with uniformity or suppress edge effects or low-order modes in the 
all-sky photometric solutions.

@@

\textbf{Joinable tables --}
Any tables for the unique and separate data products should be joinable to the 
data products for the Main Survey, when possible.
This is a requirement (Section~\ref{ssec:req_dp}).

\textbf{Survey property maps --}
As for the tesslated all-sky coadded images made from Main Survey data, survey property maps
should be made individually as part of any Special Programs that generates
tesselated coadded images. This is \emph{not} a requirement (Section~\ref{ssec:req_dp}).

@@

\item \textbf{Possible but not necessary (secondary science goals):}
a time-domain mini-survey (or DDF) has a secondary science goal of detecting
precursor outbursts for transients, which requires coadding images
within windows of days, weeks, and months to reach various depths.
This set of custom coadds may be considered as ``not necessary" and requiring 
User-Generated Processing.
Note that for the Main Survey data, such custom coadds are also considered 
as beyond scope and in need of User-Generated Processing.

@@

Consider a Special Program in which a special region of sky is observed with standard
visits but a a special strategy or cadence which is significantly distinct from the Main Survey
observing modes, and lasts for a limited amount of time.
For example, a short-term survey of the Magellanic Clouds, with dual primary science
goals in time-domain and static-sky science.
